---
title: "Credit_Score - Data Preperation - Correlation Strength"
output: html_document
date: "2024-11-10"
---


#Opening File 

```{r}
# Set the file path for the CSV file
file_path <- "C:/Users/jivko/Documents/Data Analytics, Big Data, and Predictive Analytics/Personal Project/Credit_Score_Multiclass/Credit_score_cleaned_data.csv"

# Read the CSV file into a dataframe
credit_score <- read.csv(file_path, header = TRUE)
```

```{r}
# Print the first few rows of the dataframe
print(head(credit_score))
```

```{r}
# Open the dataframe in R's data viewer (use this line once)
View(credit_score)
```

#Structure of dataframe
```{r}
str(credit_score)
```

#Missing Values Check
```{r}
# Check for missing values in each column
colSums(is.na(credit_score))
```

```{r}
# Check if there are any missing values in the dataframe
any(is.na(credit_score))
```

#Function for Numerical and Categorical Correlation - ANOVA - Eta Squared

```{r}
# Load necessary library
if (!require("lsr")) install.packages("lsr")  # lsr package provides the etaSquared function
library(lsr)

# Define function to calculate p-value, eta-squared, and show means for each level of categorical variable
calc_anova_eta_squared <- function(data, categorical_var, numeric_var) {
  
  # Convert categorical variable to factor if not already
  data[[categorical_var]] <- as.factor(data[[categorical_var]])
  
  # Perform ANOVA
  anova_model <- aov(data[[numeric_var]] ~ data[[categorical_var]], data = data)
  
  # Extract the p-value
  p_value <- summary(anova_model)[[1]][["Pr(>F)"]][1]
  
  # Format p-value: Display "<.0001" if p-value is very small
  formatted_p_value <- if (p_value < 0.0001) "<.0001" else format.pval(p_value, digits = 4)
  
  # Calculate eta-squared
  eta_squared <- etaSquared(anova_model)[1, "eta.sq"]
  
  # Calculate mean of numeric variable for each level of the categorical variable
  group_means <- tapply(data[[numeric_var]], data[[categorical_var]], mean)
  
  # Output results
  list(
    p_value = formatted_p_value,
    eta_squared = eta_squared,
    significance = ifelse(p_value < 0.05, "Significant", "Not Significant"),
    interpretation = ifelse(eta_squared < 0.06, "Weak correlation",
                            ifelse(eta_squared < 0.14, "Moderate correlation", "Strong correlation")),
    means = group_means  # Include the means of the numeric variable for each category
  )
}

# Example usage:
result <- calc_anova_eta_squared(credit_score, "Credit_Score", "Annual_Income")
print(result)
```

#Categorical and Numeric Attributes - ANOVA and Eta-Squared Calculation with Results Stored in a List for Categorical and Numeric Variables

```{r}
# Load necessary libraries
if (!require("progress")) install.packages("progress")
if (!require("lsr")) install.packages("lsr")  # lsr package for etaSquared function
library(progress)
library(lsr)

# Define function to calculate p-value, eta-squared, and show means for each level of categorical variable
calc_anova_eta_squared <- function(data, categorical_vars, numeric_vars) {
  
  # Initialize the progress bar
  total_combinations <- length(categorical_vars) * length(numeric_vars)
  pb <- progress_bar$new(
    total = total_combinations,
    format = "Processing [:bar] :percent ETA: :eta",
    clear = FALSE, width = 60
  )
  
  # Store results in a list
  results <- list()
  
  # Perform ANOVA for each combination of categorical and numeric variables
  for (cat_var in categorical_vars) {
    for (num_var in numeric_vars) {
      
      # Convert categorical variable to factor if not already
      data[[cat_var]] <- as.factor(data[[cat_var]])
      
      # Perform ANOVA
      anova_model <- aov(data[[num_var]] ~ data[[cat_var]], data = data)
      
      # Extract the p-value
      p_value <- summary(anova_model)[[1]][["Pr(>F)"]][1]
      
      # Format p-value: Display "<.0001" if p-value is very small
      formatted_p_value <- if (p_value < 0.0001) "<.0001" else format.pval(p_value, digits = 4)
      
      # Calculate eta-squared
      eta_squared <- etaSquared(anova_model)[1, "eta.sq"]
      
      # Calculate mean of numeric variable for each level of the categorical variable
      group_means <- tapply(data[[num_var]], data[[cat_var]], mean)
      
      # Store results
      results[[paste(cat_var, num_var, sep = "_")]] <- list(
        p_value = formatted_p_value,
        eta_squared = eta_squared,
        significance = ifelse(p_value < 0.05, "Significant", "Not Significant"),
        interpretation = ifelse(eta_squared < 0.06, "Weak correlation",
                                ifelse(eta_squared < 0.14, "Moderate correlation", "Strong correlation")),
        means = group_means  # Include the means of the numeric variable for each category
      )
      
      # Update progress bar
      pb$tick()
    }
  }
  
  return(results)
}

# Example usage:
# Assuming 'credit_score' is your dataframe and you have numeric and categorical columns.
categorical_vars <- c("Credit_Mix", "Payment_of_Min_Amount", "Payment_Behaviour", "Occupation", 
                      "Credit_Score", "Last_Loan_1", "Last_Loan_2", "Last_Loan_3", "Last_Loan_4",
                      "Last_Loan_5", "Last_Loan_6", "Last_Loan_7", "Last_Loan_8", "Last_Loan_9")
numeric_vars <- c("Age", "Annual_Income", "Monthly_Inhand_Salary", "Num_Bank_Accounts", "Num_Credit_Card", 
                  "Interest_Rate", "Num_of_Loan", "Delay_from_due_date", "Num_of_Delayed_Payment", 
                  "Changed_Credit_Limit", "Num_Credit_Inquiries", "Outstanding_Debt", "Credit_Utilization_Ratio", 
                  "Credit_History_Age", "Total_EMI_per_month", "Amount_invested_monthly", "Monthly_Balance")

# Call the function
result1 <- calc_anova_eta_squared(credit_score, categorical_vars, numeric_vars)

# Print the results (optional)
print(result1)
```



# Numeric and Numeric Correlations - RValue/Pearson Correlation
```{r}
# Load necessary library
if (!require("Hmisc")) install.packages("Hmisc")  # Hmisc package for rcorr function
library(Hmisc)

# Select only the relevant numeric attributes
numeric_data <- credit_score[, c("Age", "Annual_Income", "Monthly_Inhand_Salary", 
                                 "Num_Bank_Accounts", "Num_Credit_Card", "Interest_Rate", 
                                 "Num_of_Loan", "Delay_from_due_date", "Num_of_Delayed_Payment", 
                                 "Changed_Credit_Limit", "Num_Credit_Inquiries", 
                                 "Outstanding_Debt", "Credit_Utilization_Ratio", 
                                 "Credit_History_Age", "Total_EMI_per_month", 
                                 "Amount_invested_monthly", "Monthly_Balance")]

# Calculate Pearson correlation coefficients and p-values
cor_results <- rcorr(as.matrix(numeric_data), type = "pearson")

# Extract correlation coefficients and p-values
cor_matrix <- cor_results$r  # Pearson correlation coefficients
p_matrix <- cor_results$P    # p-values

# Interpret and store results
results2 <- list()  # Rename output list to "results2"

for (i in 1:(ncol(cor_matrix) - 1)) {
  for (j in (i + 1):ncol(cor_matrix)) {
    cor_val <- cor_matrix[i, j]
    p_val <- p_matrix[i, j]
    
    # Format p-value: show "<.0001" for very small values, otherwise use four decimals
    formatted_p_val <- if (p_val < 0.0001) "<.0001" else formatC(p_val, format = "f", digits = 4)
    
    # Interpretation based on specified thresholds
    interpretation <- if (cor_val == 1) {
      "Perfect positive correlation"
    } else if (cor_val == -1) {
      "Perfect negative correlation"
    } else if (cor_val >= 0.7) {
      "Strong positive correlation"
    } else if (cor_val >= 0.3) {
      "Moderate positive correlation"
    } else if (cor_val > 0) {
      "Weak positive correlation"
    } else if (cor_val > -0.3) {
      "Weak negative correlation"
    } else if (cor_val >= -0.7) {
      "Moderate negative correlation"
    } else {
      "Strong negative correlation"
    }
    
    # Significance check based on numeric p-value
    significance <- ifelse(p_val < 0.05, "Significant", "Not Significant")
    
    # If p-value is formatted as "<.0001", set significance to "Significant"
    if (formatted_p_val == "<.0001") {
      significance <- "Significant"
    }

    # Store results in the list
    results2[[paste(names(numeric_data)[i], names(numeric_data)[j], sep = "_")]] <- list(
      correlation_coefficient = cor_val,
      p_value = formatted_p_val,
      significance = significance,
      interpretation = interpretation
    )
  }
}

# Print the results (optional)
print(results2)
```


# Categorical and Categorical Correlations - Phi Coefficient/Cramer's V 
```{r}
# Load necessary libraries
if (!require("vcd")) install.packages("vcd")  # vcd package for Cramer's V
if (!require("stats")) install.packages("stats")  # For Chi-squared test
if (!require("lsr")) install.packages("lsr")  # For Cramer's V calculation
library(vcd)
library(lsr)

# Define function to calculate p-value, Cramer's V, and significance for categorical variables
calc_chisq_cramers_v <- function(data, categorical_vars) {
  
  # Initialize the progress bar
  total_combinations <- length(categorical_vars) * (length(categorical_vars) - 1) / 2
  pb <- progress_bar$new(
    total = total_combinations,
    format = "Processing [:bar] :percent ETA: :eta",
    clear = FALSE, width = 60
  )
  
  # Store results in a list
  results3 <- list()
  
  # Perform Chi-Square Test and Cramer's V for each combination of categorical variables
  for (i in 1:(length(categorical_vars) - 1)) {
    for (j in (i + 1):length(categorical_vars)) {
      
      cat_var1 <- categorical_vars[i]
      cat_var2 <- categorical_vars[j]
      
      # Create contingency table
      contingency_table <- table(data[[cat_var1]], data[[cat_var2]])
      
      # Perform Chi-Square Test
      chi_sq_test <- chisq.test(contingency_table)
      
      # Extract p-value
      p_value <- chi_sq_test$p.value
      
      # Format p-value: Display "<.0001" if p-value is very small
      formatted_p_value <- if (p_value < 0.0001) "<.0001" else format.pval(p_value, digits = 4)
      
      # Calculate Cramer's V
      cramers_v <- assocstats(contingency_table)$cramer
      association_strength <- if (cramers_v < 0.1) {
        "Weak association"
      } else if (cramers_v >= 0.1 && cramers_v < 0.3) {
        "Moderate association"
      } else if (cramers_v >= 0.3 && cramers_v < 0.5) {
        "Strong association"
      } else {
        "Very strong association"
      }
      
      # Store results in the list without chi-square statistic
      results3[[paste(cat_var1, cat_var2, sep = "_")]] <- list(
        p_value = formatted_p_value,
        significance = ifelse(p_value < 0.05, "Significant", "Not Significant"),
        cramers_v = cramers_v,
        association_strength = association_strength
      )
      
      # Update progress bar
      pb$tick()
    }
  }
  
  return(results3)
}

# Example usage:
# Assuming 'credit_score' is your dataframe with categorical variables
categorical_vars <- c("Credit_Mix", "Payment_of_Min_Amount", "Payment_Behaviour", "Occupation", 
                      "Credit_Score", "Last_Loan_1", "Last_Loan_2", "Last_Loan_3", "Last_Loan_4",
                      "Last_Loan_5", "Last_Loan_6", "Last_Loan_7", "Last_Loan_8", "Last_Loan_9")

# Call the function
results3 <- calc_chisq_cramers_v(credit_score, categorical_vars)

# Print the results (optional)
print(results3)
```


#Contigency Table for Categorical and Categorical Attributes
```{r}
# Load necessary library
if (!require("progress")) install.packages("progress")  # For progress bar
library(progress)

# Define function to calculate contingency tables with totals and percentages
get_contingency_tables_with_totals <- function(data, categorical_vars) {
  
  # Initialize the progress bar
  total_combinations <- length(categorical_vars) * (length(categorical_vars) - 1) / 2
  pb <- progress_bar$new(
    total = total_combinations,
    format = "Processing [:bar] :percent ETA: :eta",
    clear = FALSE, width = 60
  )
  
  # Store results in a list
  contingency_tables_with_totals <- list()
  
  # Generate contingency tables for each combination of categorical variables
  for (i in 1:(length(categorical_vars) - 1)) {
    for (j in (i + 1):length(categorical_vars)) {
      
      cat_var1 <- categorical_vars[i]
      cat_var2 <- categorical_vars[j]
      
      # Create contingency table
      contingency_table <- table(data[[cat_var1]], data[[cat_var2]])
      
      # Calculate totals
      row_totals <- margin.table(contingency_table, 1)  # Row totals
      col_totals <- margin.table(contingency_table, 2)  # Column totals
      grand_total <- sum(contingency_table)             # Overall total
      
      # Calculate percentages
      cell_percentage <- (contingency_table / grand_total) * 100
      row_percentage <- prop.table(contingency_table, 1) * 100
      col_percentage <- prop.table(contingency_table, 2) * 100
      
      # Create a list to store the table with totals and percentages
      contingency_tables_with_totals[[paste(cat_var1, cat_var2, sep = "_")]] <- list(
        contingency_table = contingency_table,
        row_totals = row_totals,
        col_totals = col_totals,
        grand_total = grand_total,
        cell_percentage = cell_percentage,
        row_percentage = row_percentage,
        col_percentage = col_percentage
      )
      
      # Update progress bar
      pb$tick()
    }
  }
  
  return(contingency_tables_with_totals)
}

# Example usage:
# Assuming 'credit_score' is your dataframe with categorical variables
categorical_vars <- c("Credit_Mix", "Payment_of_Min_Amount", "Payment_Behaviour", "Occupation", 
                      "Credit_Score", "Last_Loan_1", "Last_Loan_2", "Last_Loan_3", "Last_Loan_4",
                      "Last_Loan_5", "Last_Loan_6", "Last_Loan_7", "Last_Loan_8", "Last_Loan_9")

# Call the function
contingency_tables_with_totals <- get_contingency_tables_with_totals(credit_score, categorical_vars)

# Print the contingency tables with totals and percentages (optional)
print(contingency_tables_with_totals)
```


# Create dataframe, with replica to excel spreadhseet
```{r}
# Load necessary libraries
if (!require("progress")) install.packages("progress")
if (!require("lsr")) install.packages("lsr")  # lsr package for etaSquared function
if (!require("Hmisc")) install.packages("Hmisc")  # Hmisc package for rcorr function
if (!require("vcd")) install.packages("vcd")  # vcd package for Cramer's V
if (!require("stats")) install.packages("stats")  # For Chi-squared test

library(progress)
library(lsr)
library(Hmisc)
library(vcd)
library(stats)

# Define function for ANOVA and Eta-squared
calc_anova_eta_squared <- function(data, categorical_vars, numeric_vars) {
  total_combinations <- length(categorical_vars) * length(numeric_vars)
  pb <- progress_bar$new(total = total_combinations, format = "Processing [:bar] :percent ETA: :eta", clear = FALSE, width = 60)
  
  results <- data.frame()
  
  for (cat_var in categorical_vars) {
    for (num_var in numeric_vars) {
      data[[cat_var]] <- as.factor(data[[cat_var]])
      anova_model <- aov(data[[num_var]] ~ data[[cat_var]], data = data)
      p_value <- summary(anova_model)[[1]][["Pr(>F)"]][1]
      formatted_p_value <- if (p_value < 0.0001) "<.0001" else format.pval(p_value, digits = 4)
      eta_squared <- etaSquared(anova_model)[1, "eta.sq"]
      significance <- ifelse(p_value < 0.05, "Significant", "Not Significant")
      interpretation <- ifelse(eta_squared < 0.06, "Weak correlation", ifelse(eta_squared < 0.14, "Moderate correlation", "Strong correlation"))
      
      # Add to the results dataframe
      results <- rbind(results, data.frame(
        Attribute1 = cat_var, 
        Attribute2 = num_var, 
        p_value = formatted_p_value, 
        Significance = significance, 
        Interpretation = interpretation,
        Eta_Squared = eta_squared, 
        correlation_coefficient = NA, 
        cramers_v = NA
      ))
      pb$tick()
    }
  }
  return(results)
}

# Define function for Pearson Correlation
calc_pearson_correlation <- function(data, numeric_vars) {
  cor_results <- rcorr(as.matrix(data[, numeric_vars]), type = "pearson")
  cor_matrix <- cor_results$r
  p_matrix <- cor_results$P
  results <- data.frame()
  
  for (i in 1:(ncol(cor_matrix) - 1)) {
    for (j in (i + 1):ncol(cor_matrix)) {
      cor_val <- cor_matrix[i, j]
      p_val <- p_matrix[i, j]
      formatted_p_val <- if (p_val < 0.0001) "<.0001" else formatC(p_val, format = "f", digits = 4)
      significance <- ifelse(p_val < 0.05, "Significant", "Not Significant")
      interpretation <- if (cor_val == 1) {
        "Perfect positive correlation"
      } else if (cor_val == -1) {
        "Perfect negative correlation"
      } else if (cor_val >= 0.7) {
        "Strong positive correlation"
      } else if (cor_val >= 0.3) {
        "Moderate positive correlation"
      } else if (cor_val > 0) {
        "Weak positive correlation"
      } else if (cor_val > -0.3) {
        "Weak negative correlation"
      } else if (cor_val >= -0.7) {
        "Moderate negative correlation"
      } else {
        "Strong negative correlation"
      }
      
      # Add to the results dataframe
      results <- rbind(results, data.frame(
        Attribute1 = names(data)[i], 
        Attribute2 = names(data)[j], 
        p_value = formatted_p_val, 
        Significance = significance, 
        Interpretation = interpretation,
        Eta_Squared = NA, 
        correlation_coefficient = cor_val, 
        cramers_v = NA
      ))
    }
  }
  return(results)
}

# Define function for Chi-squared and Cramer's V
calc_chisq_cramers_v <- function(data, categorical_vars) {
  total_combinations <- length(categorical_vars) * (length(categorical_vars) - 1) / 2
  pb <- progress_bar$new(total = total_combinations, format = "Processing [:bar] :percent ETA: :eta", clear = FALSE, width = 60)
  
  results <- data.frame()
  
  for (i in 1:(length(categorical_vars) - 1)) {
    for (j in (i + 1):length(categorical_vars)) {
      cat_var1 <- categorical_vars[i]
      cat_var2 <- categorical_vars[j]
      contingency_table <- table(data[[cat_var1]], data[[cat_var2]])
      chi_sq_test <- chisq.test(contingency_table)
      p_value <- chi_sq_test$p.value
      formatted_p_value <- if (p_value < 0.0001) "<.0001" else format.pval(p_value, digits = 4)
      cramers_v <- assocstats(contingency_table)$cramer
      association_strength <- if (cramers_v < 0.1) {
        "Weak association"
      } else if (cramers_v >= 0.1 && cramers_v < 0.3) {
        "Moderate association"
      } else if (cramers_v >= 0.3 && cramers_v < 0.5) {
        "Strong association"
      } else {
        "Very strong association"
      }
      
      # Add to the results dataframe
      results <- rbind(results, data.frame(
        Attribute1 = cat_var1, 
        Attribute2 = cat_var2, 
        p_value = formatted_p_value, 
        Significance = ifelse(p_value < 0.05, "Significant", "Not Significant"),
        Interpretation = association_strength,
        Eta_Squared = NA, 
        correlation_coefficient = NA, 
        cramers_v = cramers_v
      ))
      pb$tick()
    }
  }
  return(results)
}

# Example usage
# Assuming 'credit_score' is your dataframe

categorical_vars <- c("Credit_Mix", "Payment_of_Min_Amount", "Payment_Behaviour", "Occupation", 
                      "Credit_Score", "Last_Loan_1", "Last_Loan_2", "Last_Loan_3", "Last_Loan_4",
                      "Last_Loan_5", "Last_Loan_6", "Last_Loan_7", "Last_Loan_8", "Last_Loan_9")
numeric_vars <- c("Age", "Annual_Income", "Monthly_Inhand_Salary", "Num_Bank_Accounts", "Num_Credit_Card", 
                  "Interest_Rate", "Num_of_Loan", "Delay_from_due_date", "Num_of_Delayed_Payment", 
                  "Changed_Credit_Limit", "Num_Credit_Inquiries", "Outstanding_Debt", "Credit_Utilization_Ratio", 
                  "Credit_History_Age", "Total_EMI_per_month", "Amount_invested_monthly", "Monthly_Balance")

# Call the functions
result1 <- calc_anova_eta_squared(credit_score, categorical_vars, numeric_vars)
result2 <- calc_pearson_correlation(credit_score, numeric_vars)
result3 <- calc_chisq_cramers_v(credit_score, categorical_vars)

# Combine all results
final_results <- rbind(result1, result2, result3)

# Print the final results
print(final_results)
```

```{r}
write.csv(final_results, "C:\\Users\\jivko\\Documents\\Data Analytics, Big Data, and Predictive Analytics\\Personal Project\\Credit_Score_Multiclass.csv", row.names = FALSE)
```

